### 2019-11-22 thesis/txt/3-data-set.tex
work on dataset section

\section{Related Data Sets}
\paragraph{TravisTorrent}

\section{Log Collection}
To collect a broad range of build logs we built the \texttt{Log Collector} using ruby.

\paragraph{Repository Sampling}
Our first task was to determine a set of repositories to query logs from.
Our \texttt{GHTorrentParser} can query the GHTorrent~\cite{gousios2013ghtorrent} data set for the most popular languages on github and the most popular repositories for a given language.
\emph{Popularity} in this case is defined as the number of watches.
Our \texttt{TravisRequester} can then check for a given repository whether it uses Travis CI.

\paragraph{Build Sampling}
The \texttt{LogCollector} uses the \texttt{TravisRequester}, our tool querying the Travis API~\cite{travisci2019apidoc}, to obtain the newest builds for a given repository.
As most builds on Travis CI are successful \mention{citation needed} we use a stratified sampling approach: \texttt{TravisRequester} saves the found builds in logs in buckets according to their state.
We encountered the following states during our data collection passes:
\begin{itemize}
	\item \todo{fill /5 mins/}~\cite{travis2009buildstatus}
\end{itemize}
\texttt{TravisRequester} saves up to a configured number of builds per state and searches through a configured number of maximum checked builds.

\paragraph{Log Sampling}
For each build the \texttt{TravisRequester} then selects a log to download.
Logs in Travis CI are not directly attributed to \texttt{build}s, but to \texttt{job}s.
A single build can consist of multiple jobs, e.g.\ building and executing tests in various different testing environments.
\texttt{TravisRequester} therefore queries each build for a job, which has the same state.
A failed build can have successful job executions, as just one failed job leads to the whole build being marked as failed.
For the selected jobs, our tool then queries the Travis API V3 and obtains the respective build log.

@@

% TODO describe somewhere else
% \section{Build Failure Reason}
% \todo{WHY? one common information developers and researches might want to extract, intentionally not focus on one structurally clearly defined retrieval   cause that would be simple for PROSE retrieval -> better comparability}

\section{Data Schema}
For our study, described in Chapter~\ref{sec:study}, we want to investigate the retrieval capabilities of three techniques. Therefore, we need a data set of in/output examples for one \texttt{BuildLogInformation}. Furthermore, Keyword search is not configured by such in/output, but search keywords surrounding the desired output. Finally, this section describes our notion of structural categories, which we use to quantify our assumptions on the needed uniformness of in/output examples.

\paragraph{Example Set}
\label{sec:example-set}
Our data set is organized in example sets. One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
Each example set has an identifying \emph{save name}, a \emph{build log information} that is desired to be extracted by this configuration, and a list of \emph{in / output examples}.

\paragraph{In / Output Example}
In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
In our data set they always consist of an \emph{input path}, linking to the build log file which is the input for this example.
The \emph{output} is a substring of that build log, the textual representation of the targeted build log information within the input build log.

\paragraph{Keywords}
The keyword search is configured by a list of keywords which are searched in the document with plain text search.
In order to compare it to the two other techniques, each in/output example is associated with a list of 1 - 3 keywords that can be found close to the textual representation of the targeted build log information.

% \paragraph{Description}
% \todo{for simple keyword search - imitate what users would search for ad-hoc}

% \todo{learning steps combining keywords -> similar to how devs would learn from reading those models}

\paragraph{Categories}
Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.
If two retrievals are at the same structural location with the build log, meaning they are surrounded by a similar markings, the fall into the same category.
For most cases, two build failure reason examples which fall into one category are outputted either within the same build step or by the same build step tool.

\section{Labeling Process}
\paragraph{Output Examples: Build Failure Reason}
We labeled the first occuring substring describing the failure.
If there were multiple errors described we took the first continuous block of error description.

The labeler read or skimmed through build log and copy out the substring describing the reason the build failed.
Whitespace or special characters were preserved, as they might be crucial for an information retrieval technique to detect the desired.
Because we are using xml to save our example sets, we also had to make sure to escape respective special characters (ampersand, less than).

\section{Validation}
\bp{some errors are not shown inside the log: we always choose an extraction, if we did not see the actual reason for the error reported, we chose the part stating that an error occured / the build failed

we are not developers of the projects and we did not analyze their travis configuration: possible that we labeled parts that presented an error, though the build did not fail because of that error

end all the points with: we tried to remedy this by (e.g.\ validations)
}
We vaildated our collected labels in two different ways. We did a second pass of labeling the build failure reason, keywords and categoies on a subset of the data and compared the results to calculate the inter-rater reliability using cohens kappa. In addition, we sent out a survey to the developers, whose commits triggered the builds within our data set and asked them whether our retrieval of the log part describing the reason the build failed was correct. In the following we describe the results of those two validation studies.

>>>>

\section{Motivation}
% For our study, described in Chapter~\ref{sec:study}, we want to investigate the retrieval capabilities of three techniques.
% Therefore, we need a data set of in/output examples for one \texttt{BuildLogInformation}.
% Furthermore, Keyword search is not configured by such in/output, but search keywords surrounding the desired output.
% Finally, this section describes our notion of structural categories, which we use to quantify our assumptions on the needed uniformness of in/output examples
In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.

\section{Related Data Sets}

\paragraph{TravisTorrent}
The \emph{TravisTorrent} data set~\cite{beller2017travistorrent} collects a broad range of metadata about builds on Travis CI.
It combines data accessible through the public Travis CI API~\cite{travisci2019apidoc}, related data from GHTorrent~\cite{gousios2013ghtorrent}, the corresponding git repository and data obtained through analysis of the build logs.
The data obtained through build logs contains the names of failing test cases, similar to the reason the build failed in \emph{LogChunks}.
However, these values are obtained through a manually developed parser, which only supports specific Ruby test runners and Java Maven or JUnit logs.
\emph{LogChunks} provides manually labeled data points on the description of the build failure reason for a much wider selection of programming languages.

\paragraph{Travis CI Build Log Data Set}
Loriot et al.~\cite{loriot2019dataset, loriot2019styler} collected a large amount of Travis CI build logs from 130 github repositories to analyze their use of the Checkstyle plugin.
They selected Maven repsitories that included the Checkstyle plugin and also used Travis CI.
Their data set only provides the plain build logs, whereas \emph{LogChunks} provides manually analyzed build logs from Travis CI.

\section{Log Collection}
In the following we describe how we select the repositories, builds and Logs for \emph{LogChunks}.
To collect the build logs we built the  \texttt{GHTorrentParser},\texttt{LogCollector} and \texttt{TravisRequester} using Ruby.

\paragraph{Repository Sampling}
First, we determine a set of repositories to query logs from.
Our \texttt{GHTorrentParser} queries the \emph{GHTorrent}~\cite{gousios2013ghtorrent} data set for the most popular languages on GitHub~\cite{github2019website}.
It then retrieves the most popular repositories for a given language.
\emph{Popularity} is defined as the number of watches.
The \texttt{TravisRequester}, our tool querying the Travis API~\cite{travisci2019apidoc}, can then check for a given repository whether it uses Travis CI.

For \emph{LogChunks} we queried GHTorrent from 01/04/2018 for the three most popular repositories of each of the 30 most popular languages.

\paragraph{Build Sampling}
The \texttt{LogCollector} uses the \texttt{TravisRequester} to obtain the newest builds for a given repository.
Wwe use a stratified sampling approach: \texttt{TravisRequester} saves the obtained builds in buckets according to their status.
We encountered the following statuses during our data collection:
\begin{itemize}
	\item created
	\item started
	\item cancelled
	\item passed
	\item errored
	\item failed
\end{itemize}
The user of \texttt{TravisRequester} configures how many builds should be checked and how many builds per status should be saved.

To sample the builds for \emph{LogChunks} we let \texttt{TravisRequester} check up to a 1000 builds per repository and kept ten for each status.
We discarded all statuses except one, namely ``failed''~\cite{travis2009buildstatus}, to make the most use of Travis' existing categorization of build status.
%Travis assigns the Status ``failed'' when the build faults in the central script phase.
%``Errored'' is assigned when the build faults during the setup or teardown phase.

\paragraph{Log Sampling}
For each build the \texttt{TravisRequester} then selects a log to download.
Travis CI attributes logs to \emph{jobs}.
A single build can consist of multiple jobs, e.g.\ building the same code version and executing tests in various different testing environments.
A failed build can have successful job executions, as just one failed job leads to the whole build being marked as failed.
\texttt{TravisRequester} queries each build for the first job, which has the same state.
For the selected jobs, the tool queries the Travis API V3 over HTTP to obtains the corresponding build log.

We manually inspected the collected build logs and had to discard logs from three repositories.
One had only one failed build, two others had empty build logs on Travis CI.
In total we collected 798 \todo{check} logs from 80 repositories.

@@

% \section{Build Failure Reason}
% \todo{WHY? one common information developers and researches might want to extract, intentionally not focus on one structurally clearly defined retrieval   cause that would be simple for PROSE retrieval -> better comparability}

\section{Data Schema}
For each repository, \emph{LogChunks} has a list of \emph{examples}.
Each example consists of:
\begin{itemize}
	\item \textbf{Input:} the relative path to the input build log.
	\item \textbf{BuildFailureReason:} the substring of the log describing the reason the build failed.
	\item \textbf{Keywords:} keywords a developer would use to search for the BuildFailureReason.
	\item \textbf{Category:} a categorization of the structural representation of the BuildFailureReason within the build log.
				The category is relative to the other examples for the same repository.
\end{itemize}
Following, this section describes the BuildFailureReason, Keywords and Category in more detail.
.

\paragraph{BuildFailureReason}
The \emph{BuildFailureReason} is the substring of the build logs that describes why the build failed.
This could be the failing test case, the description of a failed linter rule or a compiler error.
The \emph{BuildFailureReason} is one continuous string cut from the build log.
If there are multiple errors leading for the build to fail, the substring contains the first appearing continuous error descriptions.
\emph{Continuous} means that no lines reporting of normal build behavior are interrupting the error descriptions.
Wherever possible, it does \emph{not} include the log statements describing \emph{that} the build failed, but the description of \emph{why} it failed.
For a few logs we were unable to define the section detailing why the build failed, e.g. because this information was logged in another log file.
In these instances the \emph{BuildFailureReason} contains the lines describing that the build failed.

\paragraph{Keywords}
The \emph{Keywords} contain a list of one to three keywords appearing within the \emph{BuildFailureReason} or in the area around it in the build log.
We aim to select keywords a developer would use to search for the \emph{BuildFailureReason} when inspecting the build log.

\paragraph{Category}
For each repository, we assign \emph{structural categories} to the examples.
The structural category compares how the \emph{BuildFailureReasons} are represented within the build logs.
Build tools highlight their error messages with markings, e.g. starting each line with ``\texttt{ERROR}'', surrounding lines filled with special characters or additional empty log lines.
Two examples fall into the same structural categories if they are surrounded by similar markings.
For most cases, two BuildFailureReason examples which fall into one category are outputted either within the same build phase or by the same build tool.
For each repository, the structural categories are represented as integers, starting at 0 and increased with the next appearing category in chronological build order.

% \paragraph{Example Set}
% \label{sec:example-set}
% Our data set is organized in example sets. One example set always contains build log in/output examples from the same software repository or project and for one specific build log information. They are all from one repository because we investigate information retrieval techniques which are always configured to the scope of one particular software repository or project.
% Each example set has an identifying \emph{save name}, a \emph{build log information} that is desired to be extracted by this configuration, and a list of \emph{in / output examples}.

% \paragraph{In / Output Example}
% In/output examples (I/O examples) are used to configure two of the information retrieval techniques we are evaluating for this thesis, PROSE regular expression program synthesis by example and text similarity.
% In our data set they always consist of an \emph{input path}, linking to the build log file which is the input for this example.
% The \emph{output} is a substring of that build log, the textual representation of the targeted build log information within the input build log.

% \paragraph{Keywords}
% The keyword search is configured by a list of keywords which are searched in the document with plain text search.
% In order to compare it to the two other techniques, each in/output example is associated with a list of 1 - 3 keywords that can be found close to the textual representation of the targeted build log information.

% % \paragraph{Description}
% % \todo{for simple keyword search - imitate what users would search for ad-hoc}

% % \todo{learning steps combining keywords -> similar to how devs would learn from reading those models}

% \paragraph{Categories}
% Our intuition was that the ability of PROSE to be able to successfully learn a regex program depended on the structural uniformity of the provided in/output examples.
% The regular expressions need a consistent pattern in the build log string at the borders or around the textual representation of a build log information to match on for the retrieval.
% To be able to quantify this intuition in our evaluation we assigned \emph{structural categories} to each of the examples within an example set.
% If two retrievals are at the same structural location with the build log, meaning they are surrounded by a similar markings, the fall into the same category.
% For most cases, two build failure reason examples which fall into one category are outputted either within the same build step or by the same build step tool.

\section{Labeling Process}

\paragraph{BuildFailureReason}
For each repository, the labeler skimmed through the build logs and tried to identify the first occurrence of a description why the build failed.
They copied out the first continuous description as the BuildFailureReason.
They preserved whitespace and special characters, as they might be crucial to detect the targeted substring.

\paragraph{Keywords}
We presented the BuildFailureReason substring and ten lines above and below to the labeler.
Their task was to note down three strings they would put into a document search function to find this failure description.
The string should appear in or around the BuildFailureReason substring and is case-sensitive.
There are no special limitations on the string itself, especially spaces are also allowed.
Regular Expressions are not allowed.

\paragraph{Category}
To label the \emph{structural categories} we again presented the BuildFailureReason and the surrounding context to the labeler for all logs from a repository.
We asked them to assign numerical categories according to whether the BuildFailureReason substring had the same structural representation, i.e. the same surrounding or identifying characters.
The labeler should start the categories with 0 and increase as new ones appear.
For reproducibility we presented the logs in chronological build order.

\section{Validation}
We validate our collected data points in two different ways.
A different labeler performed a second pass of labeling the build failure reason, keywords and categories on a subset of the data.
We compared the results to calculate the inter-rater reliability using Cohens kappa.
In addition, we sent out a survey to the developers, whose commits triggered the builds within our data set.
We asked them whether our retrieval of the log part describing the reason the build failed was correct.
This sections describes these two validation studies.

### 2019-11-22 thesis/txt/6-discussion.tex
rework TOC with latest Feedback from Moritz, replace log screenshots
work on dataset section

\section{Choosing a Suitable Chunk Retrieval Technique}
This section unifies our study results of the chunk retrieval techniques PBE, CTS and KWS\@.
It recommends which technique to use depending on the available configuration and 
TABLE!
\paragraph{Configuration and Input}
number of categories:
one? -> pbe
two? -> CTS
more? -> KWS
available examples:
little? -> try pbe (This depends on the categories! here we might want a decision tree to represent the importance of these characteristics against each other? <- this would also anwser RQ1: which ones of the criteria we compare are the most important)
multiple -> kws
very many? -> CTS

@@

>>>>

\section{Recommendations of Suitable Techniques}
This section unifies our study results of the chunk retrieval techniques PBE, CTS and KWS\@.
It recommends which technique to use depending on the available configuration and 
TABLE!
\paragraph{Configuration and Input}
number of categories:
one? -> pbe
two? -> CTS
more? -> KWS
available examples:
little? -> try pbe (This depends on the categories! here we might want a decision tree to represent the importance of these characteristics against each other? <- this would also anwser RQ1: which ones of the criteria we compare are the most important)
multiple -> kws
very many? -> CTS

@@

some errors are not shown inside the log: we always choose an extraction, if we did not see the actual reason for the error reported, we chose the part stating that an error occured / the build failed
